---
title: "Analysis of claims"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)
library(here)
library(janitor)
library(igraph)
library(kableExtra)
library(tidylo)
library(glmmTMB)
library(broom.mixed)
library(furrr)
source("bind_weights_par.R", local = knitr::knit_global())
library(data.table)
library(dplyr)
library(readxl)
library(lme4)
library(gridExtra)
library(tibble)
library(lingtypology)
library(testthat)
library(patchwork)
library(ggplot2)
library(ggrepel)
library(grid)
library(xtable)
library(see)
library(lemon)

select <- dplyr::select
```

This notebook contains analysis of existing claims (see the main text section called "Testing 163 proposed examples of lexical elaboration"). It generates Figure 1 in the main text, and Table S1 and Figure S3 in the Supporting Material.


##Assign glottocodes for language groups

We'll create a list of glottocodes for language groups from the list of dictionaries in BILA.

Load BILA master file and collection of existing claims:

```{r loaddata, include=TRUE}
dict_path <- here("data", "biladataset", "bila_dictionaries.csv")
claims_path <- here("rawdata", "manuallycreated", "examples_of_lexical_elaboration.csv")

# we'll use WALS to assign country codes to the dictionaries in BILA
wals <- read_csv(here("rawdata", "downloaded", "wals_languages.csv"), show_col_types = FALSE) %>%
  rename(glottocode=Glottocode, countries=Country_ID) %>%
  group_by(glottocode) %>%
  summarize(countries = paste(countries, collapse = " ")) %>%
  filter(!is.na(glottocode))

dict <- read_csv(dict_path, show_col_types = FALSE) %>%
  select("glottolog_langname", "langfamily", "glottocode", "area", "affiliation") %>%
  unique() %>%
  left_join(wals %>% select(glottocode, countries), by = "glottocode") %>%
  rename(gcode = glottocode, langname = glottolog_langname) %>%
  mutate(langname_parts = str_split(langname, pattern = " "),
         affiliation_parts = str_split(affiliation, pattern = ", "),
         countries_list = map(str_split(countries, pattern = " "), ~ unique(.)))

claims_full <- read_csv(claims_path, show_col_types = FALSE) 
```

Explore and check consistency of the collection:

```{r consistencycheck, include=TRUE}
# the collection includes 122 unique concepts in 182 language or language groups
expect_equal(length(unique(claims_full$combo_name)), 122)
expect_equal(length(unique(subset(claims_full %>% filter(basic == 1))$combo_name)), 102)
expect_equal(length(unique(subset(claims_full %>% filter(basic == 0))$combo_name)), 20)

singlelangs <- claims_full %>% filter(level=="single") %>% select(gcode) %>% unique()
langroups <- claims_full %>% filter(level=="group") %>% select(gcode_rule, ref_name) %>% unique()
expect_equal(nrow(singlelangs), 136)
expect_equal(nrow(langroups), 46)
expect_equal(sum(nrow(singlelangs), nrow(langroups)), 182)

# let's check consistency, no missing values should be expected for following
expect_equal(claims_full %>% filter(is.na(langname)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(level)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(concept)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(domain)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(basic)) %>% nrow(), 0)
# no missing values in combo_name and source columns should be expected for those with lists of words
withwordlist <- claims_full %>% filter(!is.na(word_list))
expect_equal(sum(is.na(withwordlist$combo_name)), 0)
expect_equal(sum(is.na(withwordlist$source)), 0)

# 95 unique concepts have word lists, out of which 81 basic and 14 superordinates
expect_equal(length(unique(withwordlist$combo_name)), 95)
expect_equal(length(unique(subset(withwordlist %>% filter(basic == 1))$combo_name)), 81)
expect_equal(length(unique(subset(withwordlist %>% filter(basic == 0))$combo_name)), 14)

# we'd expect each unique combo_name must correspond to a unique word_list, no duplicates must be found
expect_equal(length(unique(withwordlist$combo_name)), length(unique(withwordlist$word_list)))
# no missing values should be expected for following
expect_equal(claims_full %>% filter(is.na(sign)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(reference)) %>% nrow(), 0)
expect_equal(claims_full %>% filter(is.na(scholarly)) %>% nrow(), 0)
```

We'll first assign glottocodes for language groups that don't rely on country codes:

```{r assignglottocode, include=TRUE}
claims <- claims_full %>%
  filter(level == "group") %>%
  #remove country cases for now, we'll add them later
  filter(!(gcode_rule %in% c("region_equals_to", "subregion_equals_to", "country_equals_to")))

gcode_list <- list()

for (i in seq_len(nrow(claims))) {
  if (claims$gcode_rule[i] == "langfamily_equals_to") {
    gcode_list[[i]] <- dict$gcode[dict$langfamily == claims$ref_name[i]]
  } else if (claims$gcode_rule[i] == "area_equals_to") {
    gcode_list[[i]] <- dict$gcode[dict$area == claims$ref_name[i]]
  } else if (claims$gcode_rule[i] == "langname_contains") {
    contains_language <- sapply(dict$langname_parts, function(x) claims$ref_name[i] %in% x)
    indices <- which(contains_language)
    gcode_list[[i]] <- if (length(indices) == 0) { NA } 
    else {dict$gcode[indices]}
  } else if (claims$gcode_rule[i] == "affiliation_contains") {
    contains_affiliation <- sapply(dict$affiliation_parts, function(x) claims$ref_name[i] %in% x)
    indices <- which(contains_affiliation)
    gcode_list[[i]] <- if (length(indices) == 0) {NA} 
    else {dict$gcode[indices]}
  } else {
    gcode_list[[i]] <- NULL
  }}

claims$gcode_list <- gcode_list

# let's create a list for cases that rely on country codes
countries <- read.csv2(here("rawdata", "downloaded", "un_countries.csv")) %>%
  rename(region="Region.Name", subregion="Sub.region.Name", country="Country.or.Area", code="ISO.alpha2.Code") %>%
  select(region, subregion, country, code)

country_cases <- claims_full %>%
  filter(level=="group") %>%
  filter(gcode_rule %in% c("region_equals_to", "subregion_equals_to", "country_equals_to")) %>%
  mutate(ref_name_list = map(ref_name, ~str_split(.x, ",\\s*")[[1]]))

countries_list <- list()

for (i in seq_len(nrow(country_cases))) {
  if (country_cases$gcode_rule[i] == "region_equals_to") {
    countries_list[[i]] <- countries$code[countries$region == country_cases$ref_name[i]]
  } else if (country_cases$gcode_rule[i] == "subregion_equals_to") {
    countries_list[[i]] <- countries$code[countries$subregion %in% unlist(country_cases$ref_name_list[i])]
  } else if (country_cases$gcode_rule[i] == "country_equals_to") {
    countries_list[[i]] <- countries$code[countries$country == country_cases$ref_name[i]]
  } else {
    countries_list[[i]] <- NULL
  }
}

country_cases$countries_list <- countries_list

gcode_list <- list()

for (i in seq_len(nrow(country_cases))) {
  gcode_list[[i]] <- dict$gcode[sapply(dict$countries_list, function(x) any(x %in% unlist(country_cases$countries_list[i])))]
}

gcode_list <- lapply(gcode_list, function(x) if (length(x) == 0) NA else x)

country_cases$gcode_list <- gcode_list

country_cases <- country_cases %>%
  select(-ref_name_list, -countries_list)

# add back country cases
groups_all <- bind_rows(claims, country_cases)

# we'd expect 15 language groups assigned with no glottocode because no language belonging these language groups is found in BILA
no_gcodes <- groups_all %>%
  filter(is.na(gcode_list))
expect_equal(nrow(no_gcodes), 18)
expect_equal(nrow(no_gcodes %>% select(gcode_rule, ref_name) %>% unique()), 15)
```

Remove language groups with no glottocode assigned and single languages absent from BILA.

```{r prepare dataframe, include=TRUE}
# remove language groups with no glottocode
groups_with_gcodes <- groups_all %>%
  filter(!is.na(gcode_list))

# remove single languages which are absent from BILA
singles_with_gcodes <- claims_full %>%
  filter(level=="single") %>%
  filter(gcode %in% dict$gcode) %>%
  mutate(gcode_list = str_split(gcode, " "))
expect_equal(length(unique(singles_with_gcodes$gcode)),65)

# 123 claims are excluded due to absent language and groups
bind_claims <- bind_rows(singles_with_gcodes, groups_with_gcodes) 
expect_equal(nrow(claims_full) - nrow(bind_claims), 123)
expect_equal(length(unique(bind_claims$combo_name)), 96)

# 37 claims are further excluded due to concepts where no word list was assigned
d_claims <- bind_claims %>%
  filter(!is.na(word_list)) %>%
  mutate(word_list = str_split(str_replace_all(word_list, "\\s+", ""), ",")) %>%
  select(langname, level, gcode_list, domain, basic, combo_name, word_list, sign, scholarly)
expect_equal(nrow(bind_claims) - nrow(d_claims), 37)

d_claims <- d_claims %>%
  # filter claims that appear more than once in the collection, drawn from different sources
  unique()
```

Take out dictionaries where Wiktionary filter was applied.

```{r wiktionaryfilter, include=TRUE}

wikfilter <- read_csv(here("data", "foranalyses", "wiktionary_filtered_combinations.csv"), show_col_types = FALSE)

d_claims_unnested <- d_claims %>%
  tidyr::unnest(gcode_list) %>%
  tidyr::unnest(word_list)

filter_single <- d_claims_unnested %>%
  rename(glottocode=gcode_list, word=word_list) %>%
  left_join(wikfilter, by = c("glottocode", "word")) %>%
  filter(!is.na(sum_count), level == "single") %>%
  select(langname, glottocode, combo_name) %>%
  unique()

d_claims_unnested <- d_claims_unnested %>%
  anti_join(filter_single, by=c("langname", "combo_name"))

filter_group <- d_claims_unnested %>%
  rename(glottocode=gcode_list, word=word_list) %>%
  left_join(wikfilter, by = c("glottocode", "word")) %>%
  filter(!is.na(sum_count), level == "group") %>%
  select(langname, glottocode, combo_name) %>%
  rename(gcode_list=glottocode) %>%
  unique() 

d_claims_unnested <- d_claims_unnested %>%
  anti_join(filter_group, by=c("langname", "gcode_list", "combo_name"))

# we'll use following data frames in subsequent analyses
d_claims <- d_claims_unnested %>%
  group_by(langname, level, domain, basic, combo_name, gcode_list, sign, scholarly) %>%
  mutate(word_list = list(word_list)) %>%
  ungroup()  %>%
  unique() %>%
  group_by(langname, level, domain, basic, combo_name, word_list, sign, scholarly) %>%
  mutate(gcode_list = list(gcode_list)) %>%
  ungroup() %>%
  unique()
# after wiktionary filtering, 164 claims will be further analyzed
expect_equal(nrow(d_claims), 164)

# 74 unique combinations of words will be further analyzed
d_claims_distinct <- d_claims %>%
  select(combo_name, word_list) %>%
  unique()
expect_equal(nrow(d_claims_distinct), 73)
```


##Run hierarchical model

We'll run hierarchical model to compute L^lang scores for unique combinations of words.

Load word count file:

```{r wordcount, include=TRUE}

analyses <- c("full", "robustness")

for (analysis in analyses) {

if (analysis == "full") {
  d_long_path <- here("data", "foranalyses", "d_long.csv")
  d_long <- read_csv(d_long_path, show_col_types = FALSE)
} else if (analysis == "robustness") {
  d_long_path <- here("data", "foranalyses", "d_long_robust.csv")
  d_long <- read_csv(d_long_path, show_col_types = FALSE)
}}

d_wide <- d_long %>% 
  mutate(dict = id, lang = gcode_data, langname = langname_data, family = langfamily_data, population = population_data) %>% 
  select(dict, lang, langname, family, word, count) %>% 
  group_by(dict) %>% 
  # apply smoothing
  mutate(count = count + 1) %>% 
  mutate(total = sum(count)) %>% 
  ungroup()

```

Before combining counts for combinations of words, let's see which words in word lists will be filtered out.

```{r words in stopwords, include=TRUE}

stopwords <- readLines(here("data", "foranalyses", "stopwords.txt"))
allwords <- d_wide %>%
  select(word) %>% unique() %>% pull(word)

filter_words_in_stopwords <- function(word_list) {
  word_list[word_list %in% stopwords]
}

filter_words_not_in_allwords <- function(word_list) {
  word_list[!(word_list %in% allwords) & !(word_list %in% stopwords) ]
}

filter_remaining_words <- function(word_list) {
  words_in_stopwords <- filter_words_in_stopwords(word_list)
  words_not_in_allwords <- filter_words_not_in_allwords(word_list)
  words_filtered <- c(words_in_stopwords, words_not_in_allwords)
  remaining_words <- setdiff(word_list, words_filtered)
  if (length(remaining_words) == 0) {
    return(NA)
  } else {
    return(remaining_words)
  }
}

wordsinlist <- d_claims_distinct %>%
  mutate(wordsinstop = map(word_list, filter_words_in_stopwords),
         wordsabsent = map(word_list, filter_words_not_in_allwords),
         remainingwords = map(word_list, filter_remaining_words),
         nwordsinitial = lengths(word_list),
         nwordsleft = lengths(word_list) - lengths(wordsinstop) - lengths(wordsabsent),
         nwordsfiltered = lengths(word_list) - nwordsleft) %>%
  arrange(nwordsleft) %>%
  filter(nwordsfiltered != 0)

```

Combine counts for combinations of words.

```{r combine counts, include=TRUE}
# combine counts for each combination of words 
compute_combo <- function(terms, combo) {
   result <- d_wide %>%
     filter(word %in% terms) %>%
     group_by(dict, lang, langname, family, total) %>%
     summarise(!!combo := sum(count), .groups = 'drop') %>%
     pivot_longer(cols = -c(dict, lang, langname, family, total), names_to = "word", values_to = "count") %>%
     unique()
   
   return(result)
 }

df_list <- list()

for (i in seq_len(nrow(d_claims_distinct))) {
    terms <- unlist(d_claims_distinct$word_list[i])
    combo_name <- d_claims_distinct$combo_name[i]
    df_list[[i]] <- compute_combo(terms, combo_name)
}

# combine all to single dataframe
d_combos <- bind_rows(df_list)

# take out wiktionary filtered combos
comboterm <- suppressWarnings(
  d_claims_distinct %>%
    unnest(word_list) %>%
    rename(word = word_list) %>%
    left_join(wikfilter, by = "word") %>%
    filter(!is.na(glottocode)) %>%
    select(combo_name, glottocode) %>%
    rename(word = combo_name, lang = glottocode) %>%
    distinct()
)

d_combos <- d_combos %>%
  anti_join(comboterm, by=c("lang", "word"))

# cases of words in stopwords result in 1 missing combinations, which reduces unique combinations to 72 in size. 
combos <- d_combos %>%
  rename(combo_name = word) %>%
  group_by(combo_name) %>%
  slice(1) %>%
  select(combo_name)
missing_combos <- d_claims_distinct %>%
  anti_join(combos, by="combo_name")
expect_equal(nrow(missing_combos), 1)
expect_equal(nrow(combos), 72)

# this further reduces claims to be analyzed to 163 in size.
d_claims <- d_claims %>%
  filter(!combo_name %in% missing_combos$combo_name)
expect_equal(nrow(d_claims), 163)
```

Run hierarchical model to compute L^lang score.

```{r logisticregression, include=TRUE}

weights_lr_combos <- bind_weights_par(d_combos)

for (analysis in analyses) {

if (analysis == "full") {
  
# write results for all languages in BILA
claims_lr_full <- weights_lr_combos %>%
  rename(gcode = lang) %>%
  left_join(dict %>% select(gcode, langname, langfamily, area), by = "gcode") %>%
  rename(glottocode=gcode, combo_name=word) %>%
  arrange(desc(zeta)) %>%
  write_csv(here("output", "results", "claims_lr_lang_full.csv"))

# write results for languages that appear in the collection of claims
claims_lr <- d_claims %>%
  rename(groupname=langname) %>%
  tidyr::unnest(gcode_list) %>%
  rename(glottocode=gcode_list) %>%
  left_join(claims_lr_full %>% select(glottocode, langname, combo_name, convergence, convergence_set, zeta), by=c("glottocode", "combo_name")) %>%
  unique() %>%
  arrange(desc(zeta)) %>%
  mutate(words = map_chr(word_list, str_c, collapse = ", ")) %>%
  select(-word_list) %>%
  write_csv(here("output", "results", "claims_lr_lang.csv"))
  
} else if (analysis == "robustness") {
  
# write results for all languages in BILA
claims_lr_full <- weights_lr_combos %>%
  rename(gcode = lang) %>%
  left_join(dict %>% select(gcode, langname, langfamily, area), by = "gcode") %>%
  rename(glottocode=gcode, combo_name=word) %>%
  arrange(desc(zeta)) %>%
  write_csv(here("output", "results", "claims_lr_lang_full_robust.csv"))

# write results for languages that appear in the collection of claims
claims_lr <- d_claims %>%
  rename(groupname=langname) %>%
  tidyr::unnest(gcode_list) %>%
  rename(glottocode=gcode_list) %>%
  left_join(claims_lr_full %>% select(glottocode, langname, combo_name, convergence, convergence_set, zeta), by=c("glottocode", "combo_name")) %>%
  unique() %>%
  arrange(desc(zeta)) %>%
  mutate(words = map_chr(word_list, str_c, collapse = ", ")) %>%
  select(-word_list) %>%
  write_csv(here("output", "results", "claims_lr_lang_robust.csv"))
  
}}

```


##Explore results

We'll make summary tables in three forms--expanded, default, and compressed. Expanded table contains rank scores for single languages (those in the collection as well those comprising language groups). Default table contains 191 cases of claims, with average score calculated for each language group. Compressed table is a unique combination of sign and concept, with mean rank calculated for each combination.

```{r make tables, include=TRUE}

claims_lr_full <- read_csv(here("output", "results", "claims_lr_lang_full.csv"))

# no convergence issues must be found
expect_equal(nrow(claims_lr_full %>% filter(convergence != "converged" | convergence_set != "converged")), 0)

# compute ranking
claims_nested <- claims_lr_full %>%
  group_by(combo_name) %>%
  nest()

rank_within_nest <- function(df) {
  df %>%
    mutate(rank = frank(zeta, ties.method = "random") / max(frank(zeta, ties.method = "random"))) %>%
    ungroup()
}

claims_nested <- claims_nested %>%
  mutate(data = map(data, rank_within_nest))

claims_with_rank <- claims_nested %>%
  unnest(data)

claims_lr <- read_csv(here("output", "results", "claims_lr_lang.csv"))

# combine all scores
allscores <- claims_lr %>%
  left_join(claims_with_rank %>% select(glottocode, combo_name, rank), by=c("glottocode", "combo_name")) %>%
  unique()

# expanded table
expanded <- allscores %>%
  select(sign, combo_name, scholarly, groupname, level, langname, glottocode, zeta, rank) %>%
  rename(concept=combo_name) %>%
  unique()  %>%
  mutate(concept = str_replace(concept, "combo$", ""),
         scholarly = ifelse(scholarly == 1, "scholarly", "non-scholarly")) %>%
  arrange(desc(rank)) %>%
  arrange(desc(sign))

# default table
default_first <- expanded %>%
  group_by(sign, concept, scholarly, groupname, level) %>%
  summarise(
    rank = mean(rank, na.rm = TRUE), .groups = "drop"
  ) %>%
  ungroup() %>%
  arrange(desc(rank)) %>%
  arrange(desc(sign))

# add glottlog langname to default table
default_single <- default_first %>%
  filter(level == "single") %>%
  left_join(expanded %>% select(groupname, langname), by = "groupname") %>%
  unique() %>% select(-groupname) %>%
  rename(groupname=langname)

default_group <- default_first %>% filter(level == "group")
default <- bind_rows(default_single, default_group)

# compressed table by concept
compressed <- default %>%
  group_by(sign, concept) %>%
  summarise(
    rank = mean(rank, na.rm = TRUE), .groups = "drop"
  ) %>%
   ungroup() %>%
  arrange(desc(rank)) %>%
  arrange(desc(sign))
```

Create a main figure:

```{r final plot, include=TRUE}

wordsincase <- c("snowcombo", "icecombo", "raincombo", "windcombo", "smellcombo", "tastecombo", "dancecombo")

set.seed(123) # for reproducibility when using geom_jitter

theme_font <- theme(
  text = element_text(size = 10),  # Font size for all text elements
  axis.title = element_text(size = 10),  # Font size for axis titles
  axis.text = element_text(size = 9),  # Font size for axis labels
) 
update_geom_defaults("text", list(size = 3))

langsnotinclaims_zeta <- claims_lr_full %>%
  filter(combo_name %in% wordsincase) %>%
  anti_join(claims_lr, by = c("glottocode", "combo_name")) %>%
  rename(concept=combo_name) %>%
  mutate(concept = str_replace(concept, "combo$", "")) %>%
  select(concept, langname, glottocode, zeta) %>%
  mutate(sign="not in claims")

langsinclaims_zeta <- claims_lr %>%
  filter(combo_name %in% wordsincase) %>%
  select(sign, combo_name, langname, zeta) %>%
  rename(concept=combo_name) %>%
  unique()  %>%
  mutate(concept = str_replace(concept, "combo$", ""))

allincases_zeta <- bind_rows(langsnotinclaims_zeta, langsinclaims_zeta) %>%
  filter(sign != "negative")

  snowcases <- allincases_zeta %>% 
    filter(concept == "snow") %>%
  mutate(concept = "")
zmedian <- median(snowcases$zeta)

# make alternative with 5 Inuit languages

inuit_languages <- c("Eastern Canadian Inuktitut", "Western Canadian Inuktitut", "Aleut", "Central Siberian Yupik", "North Alaskan Inupiatun")
other_snow <- c("Plains Cree", "Lakota", "Japanese", "Scots", "Navajo", "Dakota", "Ahtena", "Mi'kmaq", "North Saami", "Dena'ina", "Tibetan", "Central Carrier", "Tanacross", "Tamil", "Cebuano", "Amharic", "Sinhala")

snowcases_inuit <- snowcases %>% 
  filter( langname %in% inuit_languages)
  
snowcases_labeled<- snowcases %>% 
  filter( langname %in% inuit_languages | langname %in% other_snow) %>% 
  mutate(tcolor = if_else(langname %in% inuit_languages, "red", "black"))

plot_snow_zeta_half <- snowcases %>%
  ggplot(aes(x=1, y = zeta)) +
  geom_violinhalf(flip = TRUE) +
  xlim(-1, 4)+
  #ggplot(aes(x=zeta)) +
  geom_dotplot(binaxis = "y", stackdir = "down", dotsize = 0.2, binwidth= 0.3)  +
  geom_dotplot(data = snowcases_inuit, binaxis = "y", stackdir = "down", dotsize = 1, binwidth= 0.3, fill="red", color="red")  +
  theme_classic() +
  geom_text_repel(data = snowcases_labeled, aes(x=1,y = zeta, label = langname, color = tcolor), direction = "y", force = 1, hjust = 0,  nudge_x = 0.05, segment.color= "gray", box.padding = 0.05, size = 3) +
  scale_color_manual(values = c("black", "red")) +
  theme(legend.position="none") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  labs(x="", y=expression(L^{lang} * " for SNOW"), title = "") +
  theme_font 
plot_snow_zeta_half

default_p <- default %>%
  filter(sign == "positive" & scholarly == "scholarly") %>%
  mutate(concept = case_when(
    concept == "timeofday" ~ "time of day",
    concept == "bodyparts" ~ "body parts",
    TRUE ~ as.character(concept)),
         groupname = case_when(
      groupname == "Australian languages" ~ "Australian",
      groupname == "Southeast asian" ~ "Southeast Asian",
      groupname == "East and southeast asian" ~ "East & Southeast Asian",
      groupname == "Oceanic languages" ~ "Oceanic",
      groupname == "Suahili"  ~ "Swahili",
      TRUE ~ as.character(groupname)))

default_p_all <- default_p %>% 
  mutate(claim = paste0(toupper(concept), ", ", groupname))

labela <- default_p_all %>% filter(rank < 0.48) %>% select(claim) %>% pull(claim)

labelb <- c("SMELL, Yue Chinese", "DOCTOR, Hindi", "DANCE, Hindi",
             "TASTE, Lao", "EMOTION, Thai", "SNAKE, Arabic", "INSECTS, Navajo",
             "TREE, Venda", "EMOTION, Southeast Asian",
             "ACORN, Yana", "BODY PARTS, Aleut", "FRIEND, Russian",
             "SNOW, Eskimo-Aleut", "WIND, Oceanic",
             "ICE, Yupik", "SMELL, Thao", "TASTE, Japanese", "GLANCE, Hindi",
             "JUDGE, Swahili", "SIGHT, Yana", "ANT, Mizo", "TREE, Mi'kmaq",
            "BIRD, Papua New Guinea", "GARDENING, Oceanic", "TOGETHERNESS, Iraqw",
            "STARS, Australian", "TIME OF DAY, Hausa")

labeled <- unique(c(labela, labelb))

drop <- c("WEATHER, East & Southeast Asian", "SMELL, Arabic", "SMELL, French", "SMELL, Jehai")

default_p_unlabeled <- default_p_all %>% 
  filter( !(claim %in% labeled) )

default_p_labeled <- default_p_all %>% 
  filter( claim %in% labeled & !(claim %in% drop)) %>% 
  mutate(dot_color= if_else(claim == "SNOW, Eskimo-Aleut", "red", "black")) 

eskimosnow <- default_p_all %>% 
  filter( claim %in% "SNOW, Eskimo-Aleut" )

# make alternative with dotplot rather than jitter

default_plot_dot <- default_p %>%
  ggplot(aes(x=1, y=rank)) +
  geom_segment( aes(x=0.5, xend=1,y=0.5,yend=0.5), linetype = "dashed", size = 0.5, color = "gray") +
  geom_segment( aes(x=0.5, xend=1,y=0.95,yend=0.95), linetype = "dashed", size = 0.5, color = "gray") +
  geom_violinhalf(flip = TRUE, fill = NA) +
  xlim(0.5, 1.7) +
  geom_dotplot(binaxis = "y", stackdir = "down", dotsize = 0.2, binwidth= 0.025)  +
  geom_dotplot(data = eskimosnow, binaxis = "y", stackdir = "down", dotsize = 0.05, binwidth= 0.3, fill="red", color="red")  +
  geom_text_repel(data = default_p_labeled, aes(x=1,y = rank, label = claim, color=dot_color), direction = "y", force = 1, hjust = 0,  nudge_x = 0.05, segment.color= "gray", box.padding = 0.05, size = 3) +
  scale_color_manual(values = c("black", "red")) +
  labs(x="", y="rank", title = "") +
  theme_classic() +
  theme(legend.position="none") + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  #scale_y_continuous(limits = c(0, 1), expand = expansion(add = c(0, 0.2)))  +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), expand = expansion(add = c(0, 0.1))) +
  coord_capped_cart(left='top') +
  theme_font 
default_plot_dot
```

Plot results broken down by semantic domain

```{r final plot, include=TRUE}
# Look at results by IDS chapter
ids_features <- read_csv(here("rawdata", "downloaded", "ids_parameters.csv")) %>% 
  separate_wider_delim(ID, "-", names=c("chapter", "conceptID")) %>% 
  mutate(concept = str_replace(Name, "\\s*\\(.*\\)", "")) %>% 
  mutate(concept = str_replace(concept, "\\s*[=|,].*", ""))  %>% 
  mutate(chapter = as.numeric(chapter)) 

ids_chapters <- read_csv(here("rawdata", "downloaded", "ids_chapters.csv")) %>%  
  rename(chapter = ID, chapter_name = "Description")

default_p_ids <- default_p %>%
  left_join(ids_features, by = "concept") %>% 
  mutate(chapter = if_else(concept %in% c("rock", "soil", "lava", "stars", "wind"), 1, chapter)) %>% 
  mutate(chapter = if_else(concept %in% c("reindeer", "terrestrialmammals", "insects", "kangaroo", "crustaceans", "mollusc", "emus", "chrysalis", "locust", "eels"), 3, chapter)) %>% 
  mutate(chapter = if_else(concept %in% c("tree", "pandanus", "breadfruit", "taro", "corn", "shrub", "banana", "gardening", "fruit"), 8, chapter)) %>% 
  left_join(ids_chapters, by = "chapter") %>% 
  mutate(chapter_name = case_when(
    chapter_name == "Animals" ~ "Animals",
    chapter_name == "Agriculture and vegetation" ~ "Plants",
    chapter_name == "The physical world" ~ "Physical world",
    TRUE ~ "Other"
  )) %>%
  group_by(chapter_name, concept) %>%
  summarise(
    rank = mean(rank, na.rm = TRUE), .groups = "drop"
  ) %>%
  ungroup() %>%
  arrange(desc(rank)) %>% 
  mutate(concept = toupper(concept))  %>% 
  mutate(concept = case_when(
    concept == "BODYPARTS" ~ "BODY PARTS",
    concept == "TIMEOFDAY" ~ "TIME OF DAY",
    #concept == "INSECTS" ~ "INSECT",
    concept == "EMUS" ~ "EMU",
    concept == "EELS" ~ "EEL",
    TRUE ~ concept
  )) 
  
ids_chapter_counts <- default_p_ids %>% 
  select(concept, chapter_name) %>% 
  unique() %>% 
  group_by(chapter_name) %>% 
  summarize(count = n())

default_p_ids$chapter_name <- as.character(default_p_ids$chapter_name)
default_p_ids$chapter_name <- factor(default_p_ids$chapter_name, levels=c("Physical world", "Animals", "Plants", "Other"))

concept_labels_drop <- c("TERRESTRIALMAMMALS", "OBLIGATION", "TOGETHERNESS", "CRUSTACEANS")

default_p_ids_labeled <- default_p_ids%>% 
  filter( !( concept %in% concept_labels_drop) ) 

set.seed(0) # for reproducibility when using geom_jitter
compressed_plot_ids <- default_p_ids %>%
  ggplot(aes(x=chapter_name, y=rank)) +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_segment( aes(x=0.5, xend=4,y=0.5,yend=0.5), linetype = "dashed", size = 0.5, color = "grey") +
  geom_segment( aes(x=0.5, xend=4,y=0.95,yend=0.95), linetype = "dashed", size = 0.5, color = "grey") +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_dotplot(binaxis = "y", stackdir = "down", dotsize = 0.4, binwidth= 0.025)  +
  #geom_text_repel(aes(x=chapter_name,y = rank, label = concept), direction = "y", force = 5, hjust = 0,  nudge_x = 0.05, segment.color= "gray", box.padding = 0, size = 3) +
  geom_text_repel(data = default_p_ids_labeled, aes(x=chapter_name,y = rank, label = concept), direction = "y", force = 1, hjust = 0,  nudge_x = 0.05, segment.color= "gray", box.padding = 0.05, size = 3) +
  labs(x="", y="rank", title = "") +
  coord_cartesian(xlim = c(1.2,4), clip = "off") +
  theme_classic() +
  theme(legend.position="none") + 
  theme(axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  theme(axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), expand = expansion(add = c(0, 0.1)))  +
  coord_capped_cart(xlim = c(1.2, 4), left='top') +
  theme_font

compressed_plot_ids

plot_width <- c(0.25,0.27,0.48) 
combined_plots <- plot_snow_zeta_half + default_plot_dot + compressed_plot_ids +
  plot_layout(widths = plot_width) + plot_annotation(tag_levels = 'a',  tag_suffix = ')')

# Save the combined plot as a PDF
ggsave(filename = here::here("output", "figures", "main", "claims.pdf"), plot = combined_plots, device = "pdf", width = 15, height = 7)

```

Run permutation test to compare ranks for "other" chapter with the remaining three chapters (here called "nature")

```{r run permutation, include= TRUE}

permtest_data <- default_p_ids %>% 
  mutate(category = if_else(chapter_name == "Other", "non_nature", "nature")) %>% 
  select(category, rank)

cat_diff <- function(d) {
  d %>% group_by(category) %>% 
    summarize(meanrank = mean(rank))
}

perm_diff <- function(d) {
  cd <- d %>% 
    mutate(category = sample(category)) %>% 
    cat_diff()
  return (cd$meanrank[1] - cd$meanrank[2])
}
perms <- cross_join(tibble(i=1:10000), permtest_data) %>% 
  nest(d = c(category, rank)) %>% 
  mutate(cat_diff = map_dbl(d, perm_diff))
thresh_95 <- quantile(perms$cat_diff, probs = c(0.95))

attested_means <- cat_diff(permtest_data)
attested_diff <- attested_means$meanrank[1] - attested_means$meanrank[2] 
f <- ecdf(perms$cat_diff)
p_val = 1 - f( attested_diff )

print(paste0("attested = ", as.character(round(attested_diff, 3)), ", 95th percentile = ", as.character(round(thresh_95, 3)), ", p val = ", as.character(round(p_val, 3))))

```

##Supplementary tables and figures:

Create a table on a sample of existing claims.

```{r createtable existing claims, include=TRUE}
selected_rows <- c(1, 10, 15, 16, 19, 38, 46, 51, 53, 192, 74, 102, 107, 26, 
                   116, 118, 124, 133, 149, 140, 143, 170, 198, 251, 255, 273, 284, 292, 326, 327)

claimssample <- claims_full %>%
  arrange(combo_name) %>%
  mutate(gcode = ifelse(is.na(gcode), gcode_rule, gcode)) %>%
  select(langname, level, gcode, concept, word_list, sign, reference) %>%
  filter(!is.na(word_list)) %>%
  rename("Language/Group"=langname, Level=level, Glottocode=gcode, Concept=concept, "Word list"=word_list, Sign=sign, Reference=reference) %>%
  arrange(Concept) %>%
  slice(selected_rows)

latex_table <- xtable(claimssample)
print(latex_table, file = here("output", "tables", "claimstable.tex"), include.rownames=FALSE)
```

We make similar tables for L^lang scores computed on robustness set.

```{r make tables, include=TRUE}
claims_lr_full_robust <- read_csv(here("output", "results", "claims_lr_lang_full_robust.csv"), show_col_types = FALSE)

# no convergence issue must be found
expect_equal(length(unique(subset(claims_lr_full_robust %>% filter(convergence != "converged" | convergence_set != "converged"))$combo_name)), 0)

# compute ranking
claims_nested_robust <- claims_lr_full_robust %>%
  group_by(combo_name) %>%
  nest()

claims_nested_robust <- claims_nested_robust %>%
  mutate(data = map(data, rank_within_nest))

claims_with_rank_robust <- claims_nested_robust %>%
  unnest(data)

claims_lr_robust <- read_csv(here("output", "results", "claims_lr_lang_robust.csv"), show_col_types = FALSE)

# combine all scores
allscores_robust <- claims_lr_robust %>%
  left_join(claims_with_rank_robust %>% select(glottocode, combo_name, rank), by=c("glottocode", "combo_name")) %>%
  unique()

# expanded table
expanded_robust <- allscores_robust %>%
  select(sign, combo_name, scholarly, groupname, level, langname, zeta, rank) %>%
  rename(concept=combo_name) %>%
  unique()  %>%
  mutate(concept = str_replace(concept, "combo$", ""),
         scholarly = ifelse(scholarly == 1, "scholarly", "non-scholarly")) %>%
  arrange(desc(rank)) %>%
  arrange(desc(zeta)) %>%
  arrange(desc(sign))

# default table
default_first <- expanded_robust %>%
  group_by(sign, concept, scholarly, groupname, level) %>%
  summarise(
    rank = mean(rank, na.rm = TRUE), .groups = "drop"
  ) %>%
  ungroup() %>%
  arrange(desc(rank)) %>%
  arrange(desc(sign))

# add glottlog langname to default table
default_single <- default_first %>%
  filter(level == "single") %>%
  left_join(expanded %>% select(groupname, langname), by = "groupname") %>%
  unique() %>% select(-groupname) %>%
  rename(groupname=langname)

default_group <- default_first %>% filter(level == "group")
default_robust <- bind_rows(default_single, default_group)

# compressed table by concept
compressed_robust <- default_robust %>%
  group_by(sign, concept) %>%
  summarise(
    rank = mean(rank, na.rm = TRUE), .groups="drop"
  ) %>%
  ungroup() %>%
  arrange(desc(rank))
```

Create a figure on the full set:

```{r plot with labels, include=TRUE}

default_f <- default %>%
  mutate(concept = case_when(
      concept == "nonportableartefacts" ~ "nonportable art.",
      concept == "landscapefeatures" ~ "landscape",
      concept == "mechanicalartifacts" ~ "mechanical art.",
      concept == "terrestrialmammals" ~ "mammals",
      concept == "timeofday" ~ "time of day",
      TRUE ~ as.character(concept)),
    groupname = case_when(
      groupname == "Australian languages" ~ "Australian",
      groupname == "Southeast asian" ~ "Southeast Asian",
      groupname == "East and southeast asian" ~ "East & Southeast Asian",
      groupname == "Oceanic languages" ~ "Oceanic",
      TRUE ~ as.character(groupname)),
    claim = paste0(toupper(concept), ", ", groupname))

labelsb <- default_f %>% filter(rank <= 0.6) %>% select(claim) %>% pull(claim)
labelsc <- default_f %>% select(claim) %>% slice(1:20) %>% pull(claim)
labelsd <- c("SOIL, Venda", "RICE, East & Southeast Asian", "SHRUB, Venda", "TREE, Lau", "ANT, Lushei", "DEER, Lushei", "FISH, X창r창c첫첫", "SMELL, Paiwan", "HONEY, Arabic", "RAIN, East & Southeast Asian",
             "BEER, German", "WIND, Oceanic", "INSECTS, East & Southeast Asian", "STARS, Australian", "FISH, East & Southeast Asian", "SMELL, Khmer", "SAND, Australian", "RICE, Asian",
             "HORSE, Portuguese", "LOCUST, Arabic")

labels <- c(labelsb, labelsc, labelsd)

default_labeled <- default_f %>%
  filter(claim %in% labels)

set.seed(0) # for reproducibility when using geom_jitter
default_plot <- default_f %>%
  ggplot(aes(x = 1, y = rank)) +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_segment( aes(x=0.5, xend=4,y=0.5,yend=0.5), linetype = "dashed", size = 0.5, color = "grey") +
  geom_segment( aes(x=0.5, xend=4,y=0.95,yend=0.95), linetype = "dashed", size = 0.5, color = "grey") +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_dotplot(binaxis = "y", stackdir = "down", dotsize = 0.2, binwidth = 0.025) +
  geom_text_repel(data = default_labeled, aes(x=1, y = rank, label = claim), direction = "y", force = 1, hjust = 0, nudge_x = 0.05, segment.color= "gray", box.padding = 0.05, size = 3) +
  labs(x = "", y = "rank", title = "") +
  coord_cartesian(xlim = c(1.2, 3), clip = "off") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), expand = expansion(add = c(0, 0.1))) +
  coord_capped_cart(xlim = c(0.5, 1.5), left = 'top') +
  theme_font

default_plot

```

Create a similar figure on robustness set and combine plots:

```{r plots, include=TRUE}

default_robust <- default_robust %>%
  filter(rank != "NaN")  %>%
  mutate(concept = case_when(
      concept == "nonportableartefacts" ~ "nonportable art.",
      concept == "landscapefeatures" ~ "landscape",
      concept == "mechanicalartifacts" ~ "mechanical art.",
      concept == "terrestrialmammals" ~ "mammals",
      TRUE ~ as.character(concept)),
    groupname = case_when(
      groupname == "Australian languages" ~ "Australian",
      groupname == "Southeast asian" ~ "Southeast Asian",
      groupname == "East and southeast asian" ~ "East & Southeast Asian",
      groupname == "Oceanic languages" ~ "Oceanic",
      TRUE ~ as.character(groupname)),
    claim = paste0(toupper(concept), ", ", groupname))

labelsb <- default_robust %>% filter(rank <= 0.7) %>% select(claim) %>% pull(claim)
labelsc <- default_robust %>% select(claim) %>% slice(1:20) %>% pull(claim)
labelse <- c("COCONUT, Fijian", "FISH, Carolinian", "TREE, Niuean", "BASKET, Lushei")

labels <- c(labelsb, labelsc, labelsd, labelse)

default_robust_labeled <- default_robust %>%
  filter(claim %in% labels)

set.seed(0) # for reproducibility when using geom_jitter
default_robust_plot <- default_robust %>%
  ggplot(aes(x = 1, y = rank)) +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_segment( aes(x=0.5, xend=4,y=0.5,yend=0.5), linetype = "dashed", size = 0.5, color = "grey") +
  geom_segment( aes(x=0.5, xend=4,y=0.95,yend=0.95), linetype = "dashed", size = 0.5, color = "grey") +
  geom_violinhalf(flip = TRUE, fill = NA) +
  geom_dotplot(binaxis = "y", stackdir = "down", dotsize = 0.2, binwidth = 0.025) +
  #geom_text_repel(aes(x=chapter_name,y = rank, label = concept), direction = "y", force = 5, hjust = 0, nudge_x = 0.05, segment.color= "gray", box.padding = 0, size = 3) +
  geom_text_repel(data = default_robust_labeled, aes(x=1, y = rank, label = claim), direction = "y", force = 1, hjust = 0, nudge_x = 0.05, segment.color= "gray", box.padding = 0.05, size = 3) +
  labs(x = "", y = "rank", title = "") +
  coord_cartesian(xlim = c(1.2, 3), clip = "off") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.line.x = element_blank()) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1), expand = expansion(add = c(0, 0.1))) +
  coord_capped_cart(xlim = c(0.5, 1.5), left = 'top') +
  theme_font

default_robust_plot

plot_width <- c(0.5,0.5) 
combined_plots <- default_plot + default_robust_plot +
  plot_layout(widths = plot_width) + plot_annotation(tag_levels = 'a',  tag_suffix = ')')

# Save the combined plot as a PDF
ggsave(filename = here::here("output", "figures", "supplementary", "claimsfull.pdf"), plot = combined_plots, device = "pdf", width = 15, height = 9)
```
