---
title: "Plot related lexemes against token counts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)
library(here)
library(testthat)
library(ggplot2)
library(brms)
library(gridExtra)
library(maps)

select <- dplyr::select

```

This notebook explores the relationship between manual counts of related lexemes and the unigram frequencies for RAIN, SMELL, and TASTE in a selected subset of 64 non-HathiTrust dictionaries. It generates Figure S2 in the Supporting Material.

First we load dictionary lexeme classification data.

```{r dictionary related_lexemes, include=TRUE}

# load manual counts of dictionary related_lexemes
entrycounts <- read_csv(here("rawdata", "manuallycreated", "lexeme_classification.csv"))

# explore classification
classes <- entrycounts %>%
  group_by(word, classification) %>%
  summarise(counts=n()) %>%
  ungroup() %>%
  pivot_wider(names_from = "classification", values_from = "counts")  %>%
  mutate(total = related + unrelated + example + other) %>%
  mutate(related = round(related / total, 2),
         unrelated = round(unrelated/total, 2),
         example = round(example/total, 2),
         other = round(other/total, 2)) %>%
  arrange(desc(related))

grouped_entrycounts <- entrycounts %>%
  group_by(id, word) %>%
  summarise(
    related_lexemes = sum(classification == "related"),
    unrelated_lexemes = sum(classification %in% c("unrelated", "example", "other"))
  )

grouped <- entrycounts %>%
  group_by(id, word) %>%
  summarise(
    related = sum(classification == "related"),
    unrelated= sum(classification == "unrelated"),
    exampe = sum(classification == "example"),
    other = sum(classification == "other")
  )
```

Load word count file:

```{r load data, include=TRUE}
wordcount <- read_csv(here("data", "foranalyses", "wordcount.csv"))
dict <- read_csv(here("data", "foranalyses", "bila_dictionaries_withenv.csv"))
```

Gather counts for terms in interest. We'll compute results for groups of related words. Note that the group of words must be identical with those in analyze_cases.R. 

```{r terms in interest, include=TRUE}

rain_group <- c("rain", "raindrop", "rainfall", "rainwater", "drizzle", "mizzle", "downpour", "pelter")
smell_group <- c("smell", "odor", "scent", "effluvium", "smelling", "sniff", "snuff", "olfaction", "fragrance", "perfume", "stench")
taste_group <- c("taste", "flavor", "savor", "savoring", "gustation", "taster", "tasting", "aftertaste", "insipidity", "savoriness", "unsavoriness", "sweetness", "sourness", "acidity")

cwd <- tibble(groups=list(
  tibble(group="rain_group", word=rain_group),
  tibble(group="smell_group", word=smell_group),
  tibble(group="taste_group", word=taste_group)
  ))

wordgroups <- unnest(cwd, groups)

all_words <- wordgroups$word
all_groups <- unique(wordgroups$group)
singlewords <-  str_subset(all_groups, "_group", negate=TRUE)

all_regex <-  paste0("^(", paste(all_words, collapse="|"), ")$")
  
cases_wordcount <- wordcount %>% 
  select(id, matches(all_regex), dictsize_data) 

cases_wordcount_long <- cases_wordcount %>% 
  pivot_longer(cols = -c(id, dictsize_data), names_to = "word", values_to = "count") %>% 
  right_join(wordgroups, by = "word") 

cases_wordcount_grouped_long <- cases_wordcount_long %>% 
  group_by(id, dictsize_data, group) %>% 
  summarize(
    count = case_when(
      any(!is.na(count)) ~ sum(count, na.rm = TRUE),
      TRUE ~ NA_real_
    )
  ) %>% 
  ungroup()

cases_wordcount <- cases_wordcount_grouped_long %>%
  left_join(dict %>% select(id, glottocode), by="id") %>%
  select(-glottocode) %>%
  pivot_wider(names_from = group, values_from = count)

# we'll express frequencies in terms of counts per 10000 tokens
countunit <- 10000

# d is the dataframe we'll use for our analyses
d <- cases_wordcount %>% 
  mutate(across(-c(id, dictsize_data), ~(.x * countunit/dictsize_data))) %>% 
  mutate(countunit= countunit) %>% 
  inner_join(dict, by="id") 

# token counts
token_counts <- d %>%
  filter(id %in% grouped_entrycounts$id) %>%
  select(id, smell_group, rain_group, taste_group, dictsize_data, glottocode, langfamily) %>%
  pivot_longer(cols = c(smell_group, rain_group, taste_group), names_to = "word", values_to = "normalised_counts") %>%
  mutate(word = case_when(
    word == "smell_group" ~ "smell",
    word == "rain_group" ~ "rain",
    word == "taste_group" ~ "taste"
  )) %>%
  left_join(grouped_entrycounts, by = c("id", "word")) %>%
  mutate(raw_counts=normalised_counts * dictsize_data / countunit) %>%
  filter(raw_counts != 0)
```

## Plot related lexeme counts agains token counts

Rain:

```{r dictionary related_lexemes, include=TRUE}
theme_font <- theme(
  text = element_text(size = 10),  # Font size for all text elements
  plot.title = element_text(size=10),
  axis.title = element_text(size = 10),  # Font size for axis titles
  axis.text = element_text(size = 9),  # Font size for axis labels
  axis.ticks.x = element_blank(), 
  axis.line.x = element_blank(),
  axis.ticks.y = element_blank(), 
  axis.line.y = element_blank()
) 

# plot dictionary related_lexemes for rain related words against token counts
rain_related_lexemes <- token_counts %>%
  filter(word == "rain",
         #related_lexemes < 100,
    !is.na(normalised_counts) & !is.na(related_lexemes)
    ) %>%
  left_join(dict %>% select(id, glottocode, maxmonth_pre), by="id")

# token counts:
cor.test(rain_related_lexemes$related_lexemes, rain_related_lexemes$raw_counts)

rain_plot <- rain_related_lexemes %>%
  ggplot(aes(x = raw_counts, y = related_lexemes)) +
  geom_jitter(size=0.5) +
  geom_smooth(method=lm, size=0.3) +
  theme_classic()   +
  theme_font +
  labs(x="unigram counts", y="related lexeme counts", title="RAIN") +
  annotate("text", x = 200, y = Inf, label = "r = 0.92***", size = 4, hjust = 1.2, vjust = 1.2)
show(rain_plot)
```

Smell:

```{r dictionary related_lexemes, include=TRUE}

# plot dictionary related_lexemes for smell related words against token counts
smell_related_lexemes <- token_counts %>%
  filter(word == "smell",
    !is.na(normalised_counts) & !is.na(related_lexemes)
    ) %>%
  left_join(dict %>% select(id, glottocode), by="id")

# token counts:
cor.test(smell_related_lexemes$related_lexemes, smell_related_lexemes$raw_counts)

smell_plot <- smell_related_lexemes %>%
  ggplot(aes(x = raw_counts, y = related_lexemes)) +
  geom_jitter(size=0.5) +
  geom_smooth(method=lm, size=0.3) +
  theme_classic() + theme(axis.title.y = element_blank()) +
  theme_font +
  labs(x="unigram counts", y="related lexeme counts", title="SMELL") +
  annotate("text", x = 80, y = Inf, label = "r = 0.88***", size = 4, hjust = 1.2, vjust = 1.2)
show(smell_plot)
```

Taste:

```{r dictionary related_lexemes, include=TRUE}
# plot dictionary related_lexemes for taste related words against token counts
taste_related_lexemes <- token_counts %>%
  filter(word == "taste",
    !is.na(normalised_counts) & !is.na(related_lexemes)
    ) %>%
  left_join(dict %>% select(id, glottocode), by="id")

# token counts:
cor.test(taste_related_lexemes$related_lexemes, taste_related_lexemes$raw_counts)

taste_plot <- taste_related_lexemes %>%
  ggplot(aes(x = raw_counts, y = related_lexemes)) +
  geom_jitter(size=0.5) +
  geom_smooth(method=lm, size=0.3) +
  theme_classic()  + theme(axis.title.y = element_blank())  +
  theme_font +
  labs(x="unigram counts", y="related lexeme counts", title="TASTE") +
  annotate("text", x = 35, y = Inf, label = "r = 0.64***", size = 4, hjust = 1.2, vjust = 1.2)
show(taste_plot)
```

Make a map and combine plots:

```{r make map, include=TRUE}
langs <- token_counts %>%
  select(id) %>% unique() %>%
  left_join(dict, by = "id")

world_map <- map_data("world")

map <- ggplot() +
  geom_map(data = world_map, map = world_map,
           aes(long, lat, map_id = region),
           color = "white", fill = "gray90", size = 0.1) +
  coord_sf(xlim = c(-150, 180), ylim = c(-50, 80)) +
  geom_point(data = langs, aes(longitude, latitude), color = "#2673B4", size = 1.5, shape = 16) +
    theme_void() +
    theme(legend.position = "none")
map

library(patchwork)
combined_row <- rain_plot + smell_plot + taste_plot + plot_layout(ncol = 3, widths = c(1, 1, 1), heights = c(1.2))

combined <- combined_row / map + plot_layout(nrow = 2, heights = c(1, 1.5)) +
  plot_annotation(tag_levels = 'a', tag_suffix = ')')
ggsave(here("output", "figures", "supplementary", "lexemes_tokens.pdf"), combined, width = 10, height = 7)

```

